A Speculative Approach to Parallelization in Particle Swarm Optimization

### 2011-09-09 masters_thesis/si_paper/paper.tex
Ok, final revisions for real this time

We experimented with five common benchmark problems defined
in~\citep{herrera-2010-test-suite}: Rastrigin, Sphere, Schwefel 2.21, Griewank,
and Bohachevsky.

@@

We find benchmark functions useful in comparing algorithms, even though they
only take on the order of 1 millisecond to evaluate instead of the 100s that we
said were required to justify our parallel architecture.  The purpose of
exploring the behavior of an algorithm on benchmark functions, whether in
serial or in parallel optimization research, is because benchmark functions are
believed to represent functions that one might encounter in the real world,
regardless of the amount of time those functions take to evaluate.  While
benchmark functions themselves take fractions of a second to evaluate and thus
have no need of the kind of parallelization we explore in this paper, they also
have no need of particle swarm optimization at all, as they can be solved
analytically.  Yet they are useful to us and to optimization researchers
generally because they stand as surrogates for the kinds of functions
practitioners are actually interested in, and they allow us to explore the
behavior of optimization algorithms on several different classes of functions
in an easy and standardized way.

To provide additional evidence that performance on benchmark functions
corresponds to performance on real-world problems, we also tested our
parallelization methods on a typical research problem, that of fitting a model
to a large quantity of data.  We generated 10,000 data points from a radial
basis function network with 10 bases and some added Gaussian noise.  We then
fit a radial basis function network to the data using PSO.  This amounted to a
30 dimensional function to optimize, with a function evaluation time on the
order of two seconds.  We refer to this function as ``the model fitting
problem.''

>>>>

We experimented with five common benchmark problems defined
in~\citep{herrera-2010-test-suite}: Rastrigin, Sphere, Schwefel 2.21, Griewank,
and Bohachevsky.  While benchmark functions take fractions of a second to
evaluate and thus have no need of parallelization, they are useful for
comparing algorithms because they stand as surrogates for the kinds of
functions practitioners are actually interested in, and they allow us to
explore the behavior of optimization algorithms in a simple and standardized
way.  To provide additional evidence that performance on benchmark functions
corresponds to performance on real-world problems, we also tested our
parallelization methods on a typical research problem, that of fitting a model
to a large quantity of data, which we refer to as ``the model fitting
problem.''

@@

The model fitting problem is formulated as follows.  We generated 10,000 data
points from a radial basis function network with 10 bases and some added
Gaussian noise.  We then fit a radial basis function network to the data using
PSO.  This amounted to a 30 dimensional function to optimize, with a function
evaluation time on the order of two seconds.
